Do we need detailed demographic data to forecast plant population responses to climate change?

### 2015-06-06 manuscript/manuscript.Rmd
updates ms with comparison results section

\begin{align}
x_{jq,t+1} &= \gamma^{G}_{j,t} + \phi^{G}_{jQ} + \beta^{G}_{j,t}x_{jq,t} + \theta^{S}_{jk}C_{k,t} \\
Y^{G}_{jq,t+1} &\sim \text{LogNormal}(x_{jq,t+1}, \tau{j}) T[0,1]
\end{align}

where $x_{jq,t}$ is the log of species' $j$ proportional cover in quadrat $q$ at time $t$ and all other parameters are as in the individual-level growth model (Eq. #). The log normal likelihood includes a truncation ($T$[0,1]) to ensure that predicted values do not exceed 100% cover.

@@

All of our analyses (model fitting and simulating) were conducted in R [@R2013]. We used the 'No-U-Turn' MCMC sampler in Stan [@stan2014]  to estimate the posterior distributions of model parameters using the package 'rstan' [@rstan2014]. We obtained posterior distributions for all model parameters from three parallel MCMC chains run for 1,000 iterations after discarding an initial 1,000 iterations. We recignize such short MCMC chains may surprise those more familiar with other MCMC samplers (i.e. JAGS or WinBUGS), but the Stan sampler is exceptionally efficient, which reduces the number of iterations needed to achieve convergence. We assessed convergence visually and made sure scale reduction factors for all parameters were less than 1.01. For the purposes of introducing stochasticity in our population models, we saved the final 1,000 iterations from each of the three MCMC chains for all parameters to be used as randomly drawn values during population simulation. Importantly, this step alleviates the need to reduce model parameters since sampling from the full parameter space in the MCMC ensures that if a parameter broadly overlaps zero, on average the effect in the population models will also be near zero.

@@

Results
-------
### Comparison of Forecast Models
We performed one-step-ahead forecasts to compare the two models (IPM vs. QBM). We did not use the random year effects for these forecasts. Thus, beyond the effect of genet size or quadrat cover, only the climate covariates could affect expect plant cover. The models had similar accuracy ($\rho$, correlation between observations and mean prediction from 100 forecasts) across all species (Table 1). However, the IPM had significantly lower overall error (MAE, mean absolute error) for two species (*H. comita* and *P. smithii*), and in no case did the QBM significantly outperform the IPM (Table 1).

@@

The IPM and QBM produced consistent predictions for *H. comita* under increased precipitation and when both precipitation and temperature were increased (Figure 3). However, the IPM predicted more modest changes than the QBM, and for a temperature increase the two models differed: the IPM predicted an increase in cover while the QBM predicted the opposite.

>>>>

\begin{align}
x_{jq,t+1} &= \gamma^{G}_{j,t} + \phi^{G}_{jQ} + \beta^{G}_{j,t}x_{jq,t} + \theta^{S}_{jk}C_{k,t} \\
Y^{G}_{jq,t+1} &\sim \text{LogNormal}(x_{jq,t+1}, \tau{j}) \text{T}[0,1]
\end{align}

where $x_{jq,t}$ is the log of species' $j$ proportional cover in quadrat $q$ at time $t$ and all other parameters are as in the individual-level growth model (Eq. #). The log normal likelihood includes a truncation (T[0,1]) to ensure that predicted values do not exceed 100% cover.

@@

All of our analyses (model fitting and simulating) were conducted in R [@R2013]. We used the 'No-U-Turn' MCMC sampler in Stan [@stan2014]  to estimate the posterior distributions of model parameters using the package 'rstan' [@rstan2014]. We obtained posterior distributions for all model parameters from three parallel MCMC chains run for 1,000 iterations after discarding an initial 1,000 iterations. We recignize such short MCMC chains may surprise those more familiar with other MCMC samplers (i.e. JAGS or WinBUGS), but the Stan sampler is exceptionally efficient, which reduces the number of iterations needed to achieve convergence. We assessed convergence visually and made sure scale reduction factors for all parameters were less than 1.01. For the purposes of including parameter uncertaintyy in our population models, we saved the final 1,000 iterations from each of the three MCMC chains for all parameters to be used as randomly drawn values during population simulation. This step alleviates the need to reduce model parameters by model selection since sampling from the full parameter space in the MCMC ensures that if a parameter broadly overlaps zero, on average the effect in the population models will also be near zero.

@@

Results
-------
### Comparison of Forecast Models
We performed one-step-ahead forecasts to compare the two models (IPM vs. QBM). We did not use the random year effects for these forecasts. Thus, beyond the effect of genet size or quadrat cover, only the climate covariates could affect expect plant cover. The IPM had significantly lower overall error (MAE, mean absolute error) for two species (*H. comata* and *P. smithii*; Table 1). In no case did the QBM significantly outperform the IPM (Table 1). As a metric of model uncertainty, we calculated the euclidean distance between the upper and lower 90% quantiles for each model's forecasts. The IPM always outperformed the QBM in this respect (Table 1), with consistently lower distances between the 90% quantiles. In general the IPM outperformed the QBM because it had (i) lower MAE for two species, (ii) statistically similar MAE for the other two species, and (iii) considerably more precise forecasts for all species.

### Vital rate sensitivities to climate (IPM)

@@

The IPM and QBM produced consistent predictions for *H. comata* under increased precipitation and when both precipitation and temperature were increased (Figure 3). However, the IPM predicted more modest changes than the QBM, and for a temperature increase the two models differed: the IPM predicted an increase in cover while the QBM predicted the opposite.
